---
title: "Merge OM and Sucrose Data to Create Final Dataset"
author: "Fionna Kennedy"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

``` {r}
#' checkErrordir(path_dir)
#' 
#' Check if directory exists
#'
#' @description Errors out if the directory does not exist and gives a descriptive error.
#' @param path_dir The directory to check.
#' @return Returns error if the path does not exist, otherwise the scripts proceeds as normal.
#' @examples
#' checkErrordir("FakeDirectory101/")
checkErrordir <- function(path_dir) {
  if(dir.exists(path_dir) == FALSE) {
    stop(paste("Directory:", path_dir, "does not exist"))
  }
}

checkFile <- function(path_file) {
  if(file.exists(path_file) == FALSE) {
    stop(paste("Directory:", path_file, "does not exist"))
  }
}

#' checkMakedir(path_dir)
#' 
#' Make Non-existent Directory
#'
#' @description If the directory doesn't exist, make the directory. If the directory does exist, continue on.
#' @param path_dir The directory to check.
checkMakedir <- function(path_dir) {
  if(dir.exists(path_dir) == FALSE) {
    dir.create(path_dir)
    message(paste("Directory created:",path_dir))
  }
}
```

```{r set-wd}
# Set the working directory
wd <- "\\\\bht2stor\\vkumar\\kumarlab-new\\Fionna\\KOMP\\Ongoing Analysis\\2026-Apr-21"
setwd(wd)
```

## Load previously Processed Suc/OM Data file

```{r load-data}
# Load OM and Sucrose data
om_data <- read.csv(file.path(wd, "OM_Processed", "AnimalsAfterBadTestsRemoved_d2.csv"))
suc_data <- read.csv(file.path(wd, "Suc_Processed", "AnimalsAfterBadTestsRemoved_d2.csv"))
```

## Align Columns

Align columns of the two datasets to prepare for merging.

```{r align-columns}
# Rename DrugDose to Dose in both datasets for consistency
if ("DrugDose" %in% names(om_data)) {
  names(om_data)[names(om_data) == "DrugDose"] <- "Dose"
}
if ("DrugDose" %in% names(suc_data)) {
  names(suc_data)[names(suc_data) == "DrugDose"] <- "Dose"
}

# Add DrugMassConsumed column to sucrose data (not applicable for sucrose, set to NA)
if (!"DrugMassConsumed" %in% names(suc_data)) {
  suc_data$DrugMassConsumed <- NA
}
```

## Combine Datasets

Combine the previously cleaned OM and sucrose datasets into a single dataframe and export it.

```{r combine-and-export}
if (!require("dplyr")) install.packages("dplyr")
library(dplyr)

# Combine datasets (bind_rows handles any remaining column mismatches by filling with NA)
final_data <- bind_rows(om_data, suc_data)

# Remove unneeded columns
cols_to_remove <- c("Expr1011",
                    paste0("d", 7:12),
                    paste0("d", 14:25),
                    "coatColor", "SucrosePreference", "OralMethPreference", "Group")
final_data <- select(final_data, -any_of(cols_to_remove))

# Remove eligible outliers: only rows where n_group_iqr == "eligible" and any relevant flag is TRUE.
# "flag_only" rows (N 5-7) are never removed, matching behavior of the individual scripts.
# Use %in% TRUE for NA-safe flag evaluation (sucrose rows have NA for outlier_flag_DrugMassConsumed).
final_data_clean <- filter(
  final_data,
  !(n_group_iqr == "eligible" &
      (outlier_flag_DrugPreferenceScore %in% TRUE |
         outlier_flag_DrugMassConsumed  %in% TRUE |
         outlier_flag_DrugVolConsumed   %in% TRUE))
)

# Verify the merge
cat("Total rows (with outliers):", nrow(final_data), "\n")
cat("Total rows (outliers removed):", nrow(final_data_clean), "\n")
cat("Rows removed:", nrow(final_data) - nrow(final_data_clean), "\n")
cat("OM rows:", nrow(om_data), "\n")
cat("Sucrose rows:", nrow(suc_data), "\n")
cat("Columns:", paste(names(final_data), collapse = ", "), "\n")

# Export both versions to the same folder with descriptive names
out_dir <- file.path(wd, "Final_Merged")
checkMakedir(out_dir)

write.csv(final_data,       file.path(out_dir, "OM_Sucrose_Final_WithOutliers.csv"),    row.names = FALSE)
write.csv(final_data_clean, file.path(out_dir, "OM_Sucrose_Final_OutliersRemoved.csv"), row.names = FALSE)
message("With-outliers dataset saved to:    ", file.path(out_dir, "OM_Sucrose_Final_WithOutliers.csv"))
message("Outliers-removed dataset saved to: ", file.path(out_dir, "OM_Sucrose_Final_OutliersRemoved.csv"))
```

