---
title: "Oral Sucrose Drinking Data Analysis from JCMS for Kumar Lab"
author: "Vivek Kumar"
date: "2023-01-26" # Corrected year and added current date
output:
  html_document:
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---
# GITHUB
https://github.com/vivekJax/JCMS_OM_Suc/ 

# Sensitization Data Analysis from JCMS for Kumar Lab

**Author:** Vivek Kumar

**Date:** 2023-01-26 (corrected year and added current date)

Madison edited to provide Fionna with more pathing stability across runs.

## Introduction

This R Notebook analyzes **Drinking of Sucrose ** data
from JCMS for the Kumar Lab. The primary objective is to compare and
contrast responses between methamphetamine and sucrose data.

## Downloading Data and Directory Structure
### Notes from Tom
> Column B DataType from all data filter for 'Sucrose' or 'Oral methamphetamine'. 
>
> Last columns AU sort for 'SPT4: 2 day water, 2 days each with increasing dose of sucrose (0.25%, 0.5%, 2%) 2-bottle choice' or AV for 'OM4: 4 days 20mg-L,  4 days 40mg-L,  4 days 80mg-L oral methamphetamine' as needed. 
>
> d1 = body weight (grams), d2 = sucrose or meth water start (grams), d3 = ctrl water start (grams), d4 is sucrose or meth water end (grams), d5 = ctrl water end (grams), d6 = days drank. d13 is a comments field, usually indicating which days are covered by the indicated period since sucrose or meth experiment start (days 1-3, days 3-5, etc.). 
>
> Columns 'Drug', 'DrugDose' and 'DoseUnits' are also needed. Aside from that you just want the usual strain and sex, with MouseID, date, birthDate, Age(wks), etc. being optional.


## Importing Libraries

```{r, echo=FALSE, warning=FALSE}
library(stringr)
library(dplyr)
#library(reshape2)
library(ggplot2)
library(readxl)
library(lubridate)
library(data.table)
library(Rmisc)
library(tidyfast)
library(Hmisc)
library(viridis)           
library(ggpubr)
library(rstatix)


```

## Madison Adding some useful functions

``` {r}
#' checkErrordir(path_dir)
#' 
#' Check if directory exists
#'
#' @description Errors out if the directory does not exist and gives a descriptive error.
#' @param path_dir The directory to check.
#' @return Returns error if the path does not exist, otherwise the scripts proceeds as normal.
#' @examples
#' checkErrordir("FakeDirectory101/")
checkErrordir <- function(path_dir) {
  if(dir.exists(path_dir) == FALSE) {
    stop(paste("Directory:", path_dir, "does not exist"))
  }
}

checkFile <- function(path_file) {
  if(file.exists(path_file) == FALSE) {
    stop(paste("Directory:", path_file, "does not exist"))
  }
}

#' checkMakedir(path_dir)
#' 
#' Make Non-existent Directory
#'
#' @description If the directory doesn't exist, make the directory. If the directory does exist, continue on.
#' @param path_dir The directory to check.
checkMakedir <- function(path_dir) {
  if(dir.exists(path_dir) == FALSE) {
    dir.create(path_dir)
    message(paste("Directory created:",path_dir))
  }
}
```

``` {r}
################# SET WORKING DIRECTORY ---- FIONNA LINE 100 SHOULD BE THE ONLY LINE YOU NEED TO EDIT
wd <- "/Users/razam/Desktop/RFix_ForFionna_MR/Data"
filename <- "TOM_MasterDataTableQuery.xlsx"
checkErrordir(wd)

################# READ INPUT FILE 
filepath <- file.path(wd,filename)
checkFile(filepath)

# Marina wanted the MouseID to maintain 0 in the front --> upload 012345 which is converted to 12345 with the statement below. It convets MouseID to int
# d1 <- fread(filename, sep = '\t')

# Read and maintain the MouseID as character
# d1 <- fread(filename, sep = '\t', colClasses = c("MouseID" = "character")) # if TSV
# d1 <- read.csv(filename, sep = ',', colClasses = c("MouseID" = "character")) # if CSV
# d1 <- setDT(d1) # makes data.frame suitable for data.table

# Load EXCEL file
# Specify Column types

d1 <- as.data.table(read_excel(filepath, sheet = "TOM_MasterDataTableQuery", guess_max = 10000), showWarnings = FALSE) #made this correction to convert to data.table.

################# CREATE A FOLDER FOR OUTPUT FILES 
#  Create an empty folder in directory that is one level higher than the input file

outputDir <- file.path(dirname(wd),"Suc_Processed_FTK")
checkMakedir(outputDir)


```

# Filtering Data

**Please note:** You can either uncomment the sucrose filter or leave it
commented out depending on your current analysis focus.

## Future Steps

-   Continue with data cleaning and exploration.
-   Perform desired statistical analyses and visualizations.
-   Document results and interpretations.

**Remember to update instructions and paths as needed.**

I hope this helps!

You should now be able to copy and paste this entire code block into
your R Markdown file editor without any separation issues.

```{r}
# Filter for methamphetamine data only
# d1 <- d1[DataType == "Oral methamphetamine"]

# Filter for sucrose data only (uncomment for sucrose analysis)
  d1 <- d1[DataType == "Sucrose"]
  # d1 <- as.data.frame(d1)

```

## Add 0% sucrose
The first two days are water in both bottles. The sucrose percentage is empty in these.
Here I add 0 to these in order to use for plotting

```{r}

# adds 0 to water bottles only
d1[Drug %in% c("Both water bottles", "Both Drinking Water", "Both Water Bottles", "Both Water bottles"), "DrugDose"] <- 0

# add percent units to 
d1[DrugDose %in% 0, "DoseUnits"] <- "Percent"


```

## Calculating Drug Consumption and Preference

This script aims to calculate different metrics pertaining to drug
consumption and preference based on provided bottle weights and dosing
information. It's designed to work with both Sucrose and Meth data while
acknowledging limitations for Sucrose calculations.

**Data Preparation:**

1.  **Column Renaming:** The script starts by renaming columns from
    their numeric labels to more descriptive names for clarity. These
    changes improve understanding of the data and the calculations
    performed.

2.  **Calculation Columns:**

    -   **Drug/Water Volume Consumed:** We subtract the end weight from
        the start weight of each bottle (drug and water) to estimate the
        volume consumed in grams.
    -   **Drug Preference Score:** This represents the ratio of drug
        volume consumed to the total volume (drug + water). It ranges
        from 0 (only water) to 1 (exclusively drug).

3.  **Drug Consumed per Weight and Day:**

    -   This calculation accounts for animal weight and testing
        duration.
    -   **For Sucrose:** This calculation might not be directly
        meaningful due to its percent weight/volume dosage. While the
        script includes it for consistency, be aware of this limitation.
    -   **For Meth:** The calculation considers the milligram per liter
        dosage and converts it to milligram per gram body weight per
        day.

**Code Breakdown:**

```{r}
# Rename Body Weight Column
colnames(d1)[names(d1) == "d1"] <- "BodyWeight"

# Bottle weights
colnames(d1)[names(d1) == "d2"] <- "DrugBottleStartWeight"
colnames(d1)[names(d1) == "d3"] <- "WaterBottleStartWeight"
colnames(d1)[names(d1) == "d4"] <- "DrugBottleEndWeight"
colnames(d1)[names(d1) == "d5"] <- "WaterBottleEndWeight"

# Length of days between weighing
colnames(d1)[names(d1) == "d6"] <- "MeasureDays"

# Calculate the volume of Drug Water Consumed
# new column for Drug Volume Consumed, units are grams of liquid
d1$DrugVolConsumed <- d1$DrugBottleStartWeight - d1$DrugBottleEndWeight
d1$WaterVolConsumed <- d1$WaterBottleStartWeight - d1$WaterBottleEndWeight

# Calculate Drug Preference
# This is a ratio ranging from 0 to 1
d1$DrugPreferenceScore <- d1$DrugVolConsumed/(d1$DrugVolConsumed + d1$WaterVolConsumed)

# Drug consumed - not needed for sucrose, but I'm leaving the calculations here
# This is normalized for the weight of the animal an dthe number of days the animal is tested
# We calculatet his by multiplying the drug dose by the volume of drug consumed. This is divided by 1000 to correct for liter to ml conversion. This is then divided by the mass of the animal and divided again by the number of days between bottle weighing.
# Note that for Meth the dose is in mg/L. For sucrose it is a percent w/v. 
# We don't really use this number for sucrose. So this is OK. 
# This unit is mg/g BW/day
# d1$DrugMassConsumed <- (d1$DrugDose * d1$DrugVolConsumed/1000)/d1$BodyWeight/d1$MeasureDays
```

## Add test numbers
### Deal with Dates of tests


```{r}
# Making the date correct, if the date is not formatted correctly, adjust this I changed "%m/%d/%Y"
# this changes based on the way receive the data. It can be mdy_hms or ymd
# d1$Date <- mdy_hms(d1$Date)  # crude fix
# d1$Date <- ymd(d1$Date) # crude fix

d1$Date <- parse_date_time(d1$Date, orders = c('mdy HMS', 'ymd')) #this should handle both formats


################# SORT BY MouseID AND TestDate
d1 <- d1[order(d1$MouseID, d1$Date), ]

################# ADD TEST NUMBER AS A NEW COLUMN
# Assign TestNumber based on ordered Date within each MouseID group
d1 <- d1[ , TestNumber := order(Date), by = MouseID]

```



## Determining Animal Count per Dose (Filtered vs. Unfiltered)

This section aims to compare the number of animals across different
doses before and after applying filters.

**Outputs:**

1.  **`d1_Filtered_file.tsv`:** This file contains the filtered `d1`
    data based on specific criteria (dose, tester, strain). Note that
    filtering excludes data based on the number of tests or animals.

    -   **Important:** Duplicated mouse data exists. Utilize tables 2
        and 3 to identify and rectify such instances.

2.  **`StrainFrequencyTable_hasBadData.csv`:** This table presents a
    frequency table for each strain across different drug doses. This
    enables identifying strains lacking sufficient testing (e.g., less
    than 13 tests).

3.  **`Number_tests.csv`:** This table lists the number of tests
    conducted for each individual mouse, grouped by mouse ID and strain.
    It aids in determining missing data for specific animals.

**Reviewing these tables is crucial for ensuring data accuracy.**

**R Code:**

```{r}
# Set working directory (adjust paths as needed)
# Madison commented out and added more sure pathing
# setwd(wd)
# setwd(file.path(outputDir))

# Write filtered d1 data
fwrite(d1, file = file.path(outputDir,"d1_Filtered_file.tsv"), sep = "\t")

# Frequency table by strain and dose (identify potential missing data)
freqTable <- dcast(d1, strainName ~ DrugDose, value.var = "DrugDose", fun.aggregate = length)
fwrite(freqTable, file.path(outputDir,"StrainFrequencyTable_hasBadData.csv"))

grouped_counts <- dcast(d1, 
                         strainName + MouseID ~ DrugDose,  # Use rowid() for unique counting
                         fun.aggregate = length,  # More appropriate for counting
                         value.var = "DrugDose")  # Specify the value column

## SORT BY strainName AND MouseID
grouped_counts <- grouped_counts[order(strainName, MouseID)]

fwrite(grouped_counts, file = file.path(outputDir,"PerMouseID_Number_tests.csv"), sep = ",")




```

#   ------------------------------------------------------------------------------
### Filter and Write Data Excluding "Do Not Use" Entries
### ------------------------------------------------------------------------------
### Purpose:
### - Remove rows containing "Do Not Use" in column d13, indicating unreliable data.
### - Create a new dataset `filtered_data` with the filtered results.
### ------------------------------------------------------------------------------

```{r}
# Specify the text to exclude
excluded_text <- "Do Not Use"

# Filter the data, removing rows with the excluded text
d2 <- d1[!grepl(excluded_text, d13)]

```

#   ------------------------------------------------------------------------------
### Filter Data to remove some non-KOMP strains
### ------------------------------------------------------------------------------
### Purpose:Remove the following strains
### - B6SJL-Tg(5xFAD) HEMI
### - B6SJL-Tg(5xFAD) NCAR
### - C57BL/6J
### - C57BL/6NJ
### - DBA/2J

#   ------------------------------------------------------------------------------
### Fix certain Line Names
### ------------------------------------------------------------------------------
### - Isca had two strains, merging them here

```{r}
# Specify the text to exclude
# excluded_strains <- c("B6SJL-Tg(5xFAD) HEMI", "B6SJL-Tg(5xFAD) NCAR", "C57BL/6J", "C57BL/6NJ", "DBA/2J", "DBA/1J") # I removed this for xxx reasons 
excluded_strains <- c("B6SJL-Tg(5xFAD) HEMI", "B6SJL-Tg(5xFAD) NCAR", "C57BL/6NJ", "DBA/1J")

# Filter the data, removing rows with the excluded text
d2 <- d2[!strainName %in% excluded_strains, ]

## Fix certain names of line 
d2 <- d2[strainName == 'B6NJ-Isca1tm1a+/- Cyfip2-S968F<J> Hlethal', strainName := 'B6NJ-Isca1']
d2 <- d2[strainName == 'B6NJ-Isca1tm1b+/- Cyfip2-S968F<J> H Lethal', strainName := 'B6NJ-Isca1']

```

#   ------------------------------------------------------------------------------
### Filter and Write Data for only the 20,40,80 mg/L
### ------------------------------------------------------------------------------
### Purpose:
### - Remove rows NA for dose, these are the water only days
### - Remove rows 10 for dose, these are the early optimization tests
### - Create a new dataset `filtered_data` with the filtered results.
### - Write the filtered data to a CSV file for further analysis.
### ------------------------------------------------------------------------------

```{r}
# Specify the text to exclude for Sucrose
excluded_dose <- c("1", "4", "8", "16")

# Convert Dose to factor
# d2$DrugDose <- as.factor(d2$DrugDose)

# Filter the data, removing rows with the excluded text
d2 <- d2[!DrugDose %in% excluded_dose, ]

# Construct the output file path using clear variable names
output_file_path <- file.path(outputDir, "AnimalsAfterBadTestsRemoved_d2.csv")

# Write the filtered data to the CSV file
write.csv(d2, output_file_path)

# Print a confirmation message with the output file location
print(paste0("Filtered data successfully written to:", output_file_path))

```

# Plotting all lines on one plot, mostly for QC and viewing.

```{r}
# commented out and added more sure pathing -Madison
# setwd(wd)
# setwd(file.path(outputDir))

outFileNamePDF <- file.path(outputDir,"AllStrainsSUC.pdf")

pdf(outFileNamePDF, 12,12)

#------------------------------------------------------------------------------
PlotTitle <- "Sucrose Preference (Percent)"
xaxisLabel <- "Sucrose (%)"
yaxisLabel <- "Preference"
graphTheme <- theme_classic()
colorTheme <- "Blues"
#------------------------------------------------------------------------------
p1 <- ggplot(d2, aes(x = as.factor(DrugDose), y = DrugPreferenceScore, fill=as.factor(DrugDose))) +
      geom_rect(data = subset(d2, strainName == "B6NJ-Cyfip2 S968F<J> (GET204)"), colour = "red", fill = NA, xmin = -Inf,xmax = Inf, ymin = -Inf,ymax = Inf) + 
      geom_boxplot(outlier.shape = NA, color="gray10") +
      geom_point(shape = 1.5, color = "Gray10", position=position_jitter(0.1)) + 
      facet_wrap(~strainName, labeller = labeller(strainName = label_wrap_gen(width = 25))) +
      scale_fill_brewer(palette = colorTheme) + 
      graphTheme + 
      labs(title=PlotTitle,x= xaxisLabel, y = yaxisLabel) +
      theme(legend.position="top")

print(p1)


#------------------------------------------------------------------------------
PlotTitle <- "Sucrose Preference (Percent)"
xaxisLabel <- "TestNumber"
yaxisLabel <- "Preference"
#------------------------------------------------------------------------------

p1 <- ggplot(data = d2, aes(x = as.factor(TestNumber), y = DrugPreferenceScore, color = as.factor(DrugDose), group = c(MouseID))) +
      geom_rect(data = subset(d2, strainName == "B6NJ-Cyfip2 S968F<J> (GET204)"), colour = "red", fill = NA, xmin = -Inf,xmax = Inf, ymin = -Inf,ymax = Inf) + 
      geom_line() +
      geom_point(shape = 1.5) + 
      facet_wrap(~strainName, labeller = labeller(strainName = label_wrap_gen(width = 25))) +
      theme_classic() + 
      labs(title=PlotTitle, x=xaxisLabel, y = yaxisLabel) +
      theme(legend.position="top")
      
print(p1)
#------------------------------------------------------------------------------


#------------------------------------------------------------------------------
PlotTitle <- "Sucrose Preference (Percent)"
xaxisLabel <- "Percent Sucrose"
yaxisLabel <- "Preference"
#------------------------------------------------------------------------------

p1 <- ggplot(data = d2, aes(x = as.factor(DrugDose), y = DrugPreferenceScore, color = as.factor(DrugDose), group = c(MouseID))) +
      geom_rect(data = subset(d2, strainName == "B6NJ-Cyfip2 S968F<J> (GET204)"), colour = "red", fill = NA, xmin = -Inf,xmax = Inf, ymin = -Inf,ymax = Inf) + 
      geom_line() +
      geom_point(shape = 1.5) + 
      facet_wrap(~strainName, labeller = labeller(strainName = label_wrap_gen(width = 25))) +
      theme_classic() + 
      labs(title=PlotTitle, x=xaxisLabel, y = yaxisLabel) +
      theme(legend.position="top")
      
print(p1)
#------------------------------------------------------------------------------

#------------------------------------------------------------------------------

dev.off()

```
# QC plot for each line.
These plots are designed to find data that may need to be QCed even more.

```{r}
strainList <- levels(as.factor(d2$strainName))

# Commented otu and added more sure pathing -Madison
# setwd(wd)
# setwd(file.path(outputDir))

outFileNamePDF <- file.path(outputDir,"SucPrefIndividualStrains.pdf") 

pdf(outFileNamePDF, 6,3)

for (toPlot in strainList) {
  toPlot1 <- d2[d2$strainName %in% c(toPlot), ]
  
p1 <- ggplot(toPlot1, aes(x = as.factor(TestNumber), y = DrugPreferenceScore,color = DrugDose, group = MouseID)) +
      geom_line() + 
      geom_point(shape = 1.5) +   
      theme_classic() + 
      labs(title=stringr::str_wrap(toPlot, width=40), y = "% Pref") + 
      theme(legend.position="right") +
      scale_y_continuous(limits = c(-.5, 1.5)) +
      guides(col = guide_legend(nrow = 8)) +
      facet_grid(.~sex)

print(p1)
 
  
p1 <- ggplot(toPlot1, aes(x = as.factor(DrugDose), y = DrugPreferenceScore, color=MouseID, group = MouseID)) +
      geom_line() + 
      geom_point(shape = 1.5) +   
      theme_classic() + 
      labs(title=stringr::str_wrap(toPlot, width=40), y = "% Pref") + 
      theme(legend.position="right") +
      scale_y_continuous(limits = c(-.5, 1.5)) +
      guides(col = guide_legend(nrow = 8)) +
      facet_grid(.~sex)

print(p1)

p1 <- ggplot(toPlot1, aes(x = as.factor(DrugDose), y = DrugPreferenceScore, color=MouseID, group = MouseID)) +
      geom_line() + 
      geom_point(shape = 1.5) +   
      theme_classic() + 
      labs(title=stringr::str_wrap(toPlot, width=40), y = "% Pref") + 
      theme(legend.position="right") +
      scale_y_continuous(limits = c(-.5, 1.5)) +
      guides(col = guide_legend(nrow = 8)) +
      facet_wrap(.~MouseID) +
      theme(legend.position="none") +
      theme(axis.text = element_text( size = 5),
           axis.text.x = element_text( size = 5),
           axis.title = element_text( size = 5, face = "bold" ),
           legend.position="none",
           # The new stuff
           strip.text = element_text(size = 5))


print(p1)
 
}


dev.off()
```

# Plot each line compared to the controls
## Box plots with points

```{r}
strainList <- levels(as.factor(d2$strainName))

# Commented out and added more sure pathing -Madison
# setwd(wd)
# setwd(file.path(outputDir))

outFileNamePDF <- file.path(outputDir,"IndividualStrainPlots.pdf") 

pdf(outFileNamePDF, 4,3)

for (toPlot in strainList) {
  # if the strain is control then skip
  if(toPlot=="B6NJ-Cyfip2 S968F<J> (GET204)") next
  # extract the animal to plot
  toPlot1 <- d2[d2$strainName %in% c(toPlot, "B6NJ-Cyfip2 S968F<J> (GET204)"), ]
  
p1 <- ggplot(toPlot1, aes(x = as.factor(DrugDose), y = DrugPreferenceScore , fill=as.factor(DrugDose))) +
      geom_boxplot(outlier.shape = NA, color="gray10", width = 0.65) +
      facet_wrap(~strainName, labeller = labeller(strainName = label_wrap_gen(width = 25))) +
      scale_fill_brewer(palette="Blues") + 
      theme_classic() + 
      labs(title=stringr::str_wrap(toPlot, width=40), y = "mgMeth/gBW/day") + 
      theme(plot.title = element_text(size = 10, face = "bold")) +
      theme(legend.position="none") + 
      theme(strip.text.x = element_text(size = 8)) # decrease facet font size


      print(p1)
 
  
   p1 <- p1 +
      geom_point(shape = 1.5, color = "Gray10", position=position_jitter(0.1)) 
      
      print(p1)
}

dev.off()


```

# Plot each line compared to the controls facet by sex

```{r}
strainList <- levels(as.factor(d2$strainName))

# Commented out and addded more sure pathing -Madison
# setwd(wd)
# setwd(file.path(outputDir))

outFileNamePDF <- file.path(outputDir,"IndividualStrainPlotsSEX.pdf") 

pdf(outFileNamePDF, 4.5,6)

for (toPlot in strainList) {
  # if the strain is control then skip
  if(toPlot=="B6NJ-Cyfip2 S968F<J> (GET204)") next
  # extract the animal to plot
  toPlot1 <- d2[d2$strainName %in% c(toPlot, "B6NJ-Cyfip2 S968F<J> (GET204)"), ]
  
# removed an errant comma -Madison 
p1 <- ggplot(toPlot1, aes(x = as.factor(DrugDose), y = DrugPreferenceScore , fill=as.factor(DrugDose))) +
      geom_boxplot(outlier.shape = NA, color="gray10", width = 0.65) +
      facet_grid(sex~strainName, labeller = labeller(strainName = label_wrap_gen(width = 25))) +
      scale_fill_brewer(palette="Blues") + 
      theme_bw() + 
      labs(title=stringr::str_wrap(toPlot, width=40), y = "Sucrose Preference (%)") + 
      theme(plot.title = element_text(size = 10, face = "bold")) +
      theme(legend.position="none") + 
      theme(strip.text.x = element_text(size = 8), # decrease facet font size
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank()) 
      print(p1)
 
  
   p1 <- p1 +
      geom_point(shape = 1.5, color = "Gray10", position=position_jitter(0.1)) 
      
      print(p1)
}

dev.off()

```

# Plot all animals as a line plot, faceted by line 
## this is useful for trying to find outliers
```{r}
# Commented out and added more sure pathing -Madison
# setwd(wd)
# setwd(file.path(outputDir))

outFileNamePDF <- file.path(outputDir,"AllStrainsLine_Sucrose.pdf") 

pdf(outFileNamePDF, 8,8)

#------------------------------------------------------------------------------
PlotTitle <- "Sucrose Preference (Percent)"
xaxisLabel <- "TestNumber"
yaxisLabel <- "Preference"
#------------------------------------------------------------------------------

p1 <- ggplot(data = d2, aes(x = as.factor(DrugDose), y = DrugPreferenceScore, color = strainName, group = c(MouseID))) +
      geom_line() +
      geom_point(shape = 1.5) + 
      facet_wrap(~strainName, labeller = labeller(strainName = label_wrap_gen(width = 25))) +
      theme_classic() + 
      labs(title=PlotTitle, x=xaxisLabel, y = yaxisLabel) +
      theme(legend.position="NA")
      
print(p1)
#------------------------------------------------------------------------------

dev.off()

```


```{r}
#------------------------------------------------------------------------------
# Line Plot
PlotTitle <- ""
xaxisLabel <- "Sucrose (%)"
yaxisLabel <- "Preference"
# The errorbars overlapped, so use position_dodge to move them horizontally
pd <- position_dodge(0.1) # move them .05 to the left and right

#------------------------------------------------------------------------------

strainList <- levels(as.factor(d2$strainName))

# Commented out and added more sure pathing -Madison
# setwd(wd)
# setwd(file.path(outputDir))

outFileNamePDF <- file.path(outputDir,"IndividualStrainLinePlot_Sucrose.pdf") 

pdf(outFileNamePDF, 4,3)

for (toPlot in strainList) {
  # if the strain is control then skip
  if(toPlot=="B6NJ-Cyfip2 S968F<J> (GET204)") next
  # extract the animal to plot
  toPlot1 <- d2[d2$strainName %in% c(toPlot, "B6NJ-Cyfip2 S968F<J> (GET204)"), ]



p1 <- ggplot(toPlot1, aes(x = as.factor(DrugDose), y = DrugPreferenceScore, color = strainName, group = strainName)) +
      geom_linerange(stat = "summary", fun.data = mean_se, position=pd, color = "gray50") +
      stat_summary(fun='mean', geom='point', position=pd, shape = 19) + 
      stat_summary(fun="mean", geom='line', position=pd) +
      graphTheme + 
      labs(title = toPlot, x= xaxisLabel, y = yaxisLabel) +
      theme(legend.position="top") +
      guides(col = guide_legend(nrow = 2)) +
      theme(legend.title=element_blank()) +
      scale_color_brewer(palette = "Dark2") +
      theme(plot.title = element_text(size = 8, face = "bold")) 
print(p1)  

p1 <- p1 + facet_grid(.~ sex)
print(p1)  
}

dev.off()

#------------------------------------------------------------------------------
################ Add significance testing
# https://www.datanovia.com/en/blog/how-to-add-p-values-onto-a-grouped-ggplot-using-the-ggpubr-r-package/


outFileNamePDF <- file.path(outputDir,"IndividualStrainLinePlotPvalues_Suc.pdf") 

pdf(outFileNamePDF, 4,3)

for (toPlot in strainList) {
      tryCatch({

  # if the strain is control then skip
  if(toPlot=="B6NJ-Cyfip2 S968F<J> (GET204)") next
  # extract the animal to plot
  toPlot1 <- d2[d2$strainName %in% c(toPlot, "B6NJ-Cyfip2 S968F<J> (GET204)"), ]
  toPlot1$DrugDose <- as.numeric(toPlot1$DrugDose) # the drug dose are charcters


stat.test <- toPlot1 %>%
  group_by(DrugDose) %>%
  t_test(DrugPreferenceScore ~ strainName) %>%
  adjust_pvalue(method = "none") %>%
  add_significance("p.adj")
stat.test


p1 <- ggline(toPlot1, 
             x = "DrugDose", 
             y = "DrugPreferenceScore", 
             add = "mean_se", 
             color = "strainName", 
             position=pd, 
             palette = c("#00AFBB", "#E7B800"))
p1 <- p1 + labs(color=NULL) 
p1 <- p1 + guides(col = guide_legend(nrow = 2))

# Add p-values onto the line plots
# Remove brackets using linetype = "blank"
stat.test <- stat.test %>%
  add_xy_position(fun = "mean_se", x = "DrugDose") 

p1 <- p1 + stat_pvalue_manual(
  stat.test,  label = "{p.adj}{p.adj.signif}", 
  tip.length = 0, linetype  = "blank")
  
print(p1)
}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

}

dev.off()

#------------------------------------------------------------------------------
################ Add significance testing by sex
# https://www.datanovia.com/en/blog/how-to-add-p-values-onto-a-grouped-ggplot-using-the-ggpubr-r-package/
# https://www.datanovia.com/en/blog/add-p-values-to-ggplot-facets-with-different-scales/


outFileNamePDF <- file.path(outputDir,"IndividualStrainLinePlotSexPvalues_Suc.pdf") 

pdf(outFileNamePDF, 6,4)
for (toPlot in strainList) {
    tryCatch({
  
  # if the strain is control then skip
  if(toPlot=="B6NJ-Cyfip2 S968F<J> (GET204)") next
  # extract the animal to plot
  toPlot1 <- d2[d2$strainName %in% c(toPlot, "B6NJ-Cyfip2 S968F<J> (GET204)"), ]
  toPlot1$DrugDose <- as.numeric(toPlot1$DrugDose) # the drug dose are charcters


stat.test <- toPlot1 %>%
  group_by(DrugDose, sex) %>%
  t_test(DrugPreferenceScore ~ strainName) %>%
  adjust_pvalue(method = "none") %>%
  add_significance("p.adj")
stat.test


p1 <- ggline(toPlot1, 
             x = "DrugDose", 
             y = "DrugPreferenceScore", 
             add = "mean_se", 
             color = "strainName", 
             position=pd, 
             palette = c("#00AFBB", "#E7B800"),
             facet.by = "sex")
p1 <- p1 + labs(color=NULL) 
p1 <- p1 + guides(col = guide_legend(nrow = 2))

# Add p-values onto the line plots
# Remove brackets using linetype = "blank"
stat.test <- stat.test %>%
  add_xy_position(fun = "mean_se", x = "DrugDose") 

p1 <- p1 + stat_pvalue_manual(stat.test,  label = "{p.adj}{p.adj.signif}", tip.length = 0, linetype  = "blank")

print(p1)

}, error=function(e){cat("ERROR :",conditionMessage(e), "\n")})

}
dev.off()


```


# Check Body Weights Over Tests
> summary(d2$BodyWeight)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  11.70   19.80   22.60   22.77   25.30  266.00 
  
### SET range from 15 to 30, it should cover most of the animals

```{r}
strainList <- levels(as.factor(d2$strainName))

# setwd(wd)
# setwd(file.path(outputDir))

outFileNamePDF <- file.path(outputDir,"BodyWeightIndividualStrains.pdf") 

pdf(outFileNamePDF, 6,3)

for (toPlot in strainList) {
  toPlot1 <- d2[d2$strainName %in% c(toPlot), ]
  
p1 <- ggplot(toPlot1, aes(x = as.factor(TestNumber), y = BodyWeight , color=MouseID, group = MouseID)) +
      geom_line() + 
      geom_point(shape = 1.5) +   
      theme_classic() + 
      labs(title=stringr::str_wrap(toPlot, width=40), y = "Grams") + 
      theme(legend.position="right") +
      scale_y_continuous(limits = c(15, 35)) +
      guides(col = guide_legend(nrow = 8)) +
      facet_grid(.~sex)

print(p1)
 
 }

dev.off()

# --------------------------------  PLOT ALL MICE ON ONE PLOT
outFileNamePDF <- file.path(outputDir,"BodyWeightAllMice.pdf") 

pdf(outFileNamePDF, 12,9)

# fix one datapoint that is too high
d2[BodyWeight == 266, BodyWeight := 26.6]


p1 <- ggplot(d2, aes(x = as.factor(TestNumber), y = BodyWeight, group = MouseID)) +
      geom_line(aes(color = BodyWeight), alpha = 0.5, size = 0.5) + 
      theme_classic() + 
      labs(title="BW of All sucrose tests", y = "Grams") + 
      theme(legend.position="NA") +
      facet_grid(.~sex) +
      scale_color_viridis(option = "D")


print(p1)

dev.off()

```
